---
title: "Enterococcus Report Card Process"
format: html
message: false
warning: false
error: true
---

# Load things  

## Packages, spatial layers, etc.  

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(tbeptools)

dbasins <- read_sf(here::here("data", "geospatial", "dbasins-galb.shp"))
catchments <- read_sf(here::here("data", "geospatial", "TB_Select_Fib_Catchments_Dissolved.shp"))
swfwmd_pixels <- read_sf(here::here("data", "geospatial", "swfwmd_pixel_2_utm_m_83.shp"))
```



## Source functions  

These will eventually be incorporated into tbeptools; for now sourcing the .R files

```{r}
# to download Entero data at specified stations
source(here::here("R", "read_importwqpStationsAndStuff.R"))

# to download precip data from SWFWMD
source(here::here("R", "raindl.R"))

# to identify wet/dry fib samples  
source(here::here("R", "anlz_fibwetdry.R"))

source(here::here("R", "anlz_enteromatrix.R"))
source(here::here("R", "show_enteromatrix.R"))

source(here::here("R", "anlz_enteromap.R"))
source(here::here("R", "show_enteromap.R"))
```



# Download things  


## Enterococcus data  

```{r}
stns <- unique(catch_pixels$station)
stns[which(stns == "21FLCOSP_WQX-CENTRAL *")] <- "21FLCOSP_WQX-CENTRAL CANAL"
# define arguments


stations <- stns  # can subset if desired
entero_names <- c("Enterococci",
                  "Enterococcus")
startDate <- as.Date("2010-01-01")
endDate <- as.Date("2022-12-31")
```

```{r}
# put them together
# date format has to be mm-dd-yyyy for the url
args <- list(
    siteid = stations,
    characteristicName = entero_names,
    startDateLo = format(startDate, "%m-%d-%Y"),
    startDateHi = format(endDate, "%m-%d-%Y")
)
```

```{r}
# grab the data
fibdat <- read_importwqpStationsAndStuff(args = args)

# fibdat2 <- read_importwqpStationsAndStuff(args = args)

fibdat$date <- lubridate::ymd(fibdat$date)
```



```{r}
#| eval: false

fibdat2 <- fibdat |> 
    mutate(year = lubridate::year(date))
table(fibdat2$station, fibdat2$year)
```


## Precip data  

Actually need to loop through some years  

```{r}
# yrs <- 2018:2022
# 
# prcp_out <- list()
# 
# for(i in seq_along(yrs)){
#     yr = yrs[i]
#     prcptmp <- read_importrain(curyr = yr, catch_pixels = catch_pixels,
#                            quiet = FALSE)
#     prcp_out[[i]] <- prcptmp
# }
# 
# prcpdat <- bind_rows(prcp_out)
# 
# save(prcpdat, file = here::here("data", "precipitation", "daily_rain_stns_2018-2022.RData"))
load(here::here("data", "precipitation", "daily_rain_stns_2018-2022.RData"))
```

Fix Central Canal in the precip data station listing, so it will match the station in the fib data.  

```{r}
prcpdat <- prcpdat |> 
    mutate(station = case_when(station == "21FLCOSP_WQX-CENTRAL *" ~ "21FLCOSP_WQX-CENTRAL CANAL",
                               .default = station))
```





# Link things  

Entero data to precip data; identify wet/dry samples  

```{r}
tmpthrs <- 3
wetthrs <- .5

# define 'wet' sample as 1 inch over 2 days
test1 <- anlz_fibwetdry(fibdat, prcpdat, 
                        temporal_window = tmpthrs, 
                        wet_threshold = wetthrs)
# number of wet samples:
sum(test1$wet_sample, na.rm = TRUE)
# proportion of samples that are wet:
(prop_wet <- sum(test1$wet_sample, na.rm = TRUE)/nrow(test1))
pct_wet <- round(prop_wet*100, 1)
nsamps <- nrow(test1)
nstns <- length(unique(test1$station))

ggplot(test1, aes(x = wet_sample,
                  y = ecocci)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(subtitle = glue::glue("\n{pct_wet}% of {nsamps} samples across {nstns} stations 'wet' \n'wet' threshold {wetthrs} inches over {tmpthrs} days"))


ggplot(test1, aes(x = wet_sample,
                  y = rain_total)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(subtitle = glue::glue("'wet' threshold {wetthrs} inches over {tmpthrs} days"))

ggplot(test1, aes(x = wet_sample,
                  y = rain_total)) +
    geom_boxplot() +
    labs(subtitle = glue::glue("'wet' threshold {wetthrs} inches over {tmpthrs} days"))
```



```{r}
# define 'wet' sample as 0.5 inches in 1 day
test2 <- anlz_fibwetdry(fibdat, prcpdat, 
                        temporal_window = 1, 
                        wet_threshold = 0.2)
# number of wet samples:
sum(test2$wet_sample)
# proportion of samples that are wet:
sum(test2$wet_sample)/nrow(test2)
```


from an email 6/12/24:
"Basically, 1/2 inch over 2 days (sample day + one day prior) gets pretty similar output as 3/4 inch over 3 days (sample day + 2 days prior). Both of those options flag about 18% of samples as 'wet'. 1/2 inch over 3 days flags about 25% as wet, as does 3/4 inches over 4 days. Other combinations gave 10% or less, or 30% or more - all ranges that felt a little wrong.  

Anyway, 1/2 inch over 2 days and 3/4 inch over 3 days both "feel" pretty good, and I think there's precedent for either choice, if I'm remembering right from my other research."


```{r}
fibwetdry <- anlz_fibwetdry(fibdat, prcpdat, 
                        temporal_window = 2, 
                        wet_threshold = 0.5) |> 
    filter(!is.na(wet_sample))
# number of wet samples:
sum(fibwetdry$wet_sample, na.rm = TRUE)
# proportion of samples that are wet:
sum(fibwetdry$wet_sample, na.rm = TRUE)/nrow(fibwetdry)



# split into wet/dry frames
fibwet <- fibwetdry |> 
    filter(wet_sample == TRUE)

fibdry <- fibwetdry |> 
    filter(wet_sample == FALSE)
```



# Calculate Report Card Metrics  

make it like epchc data to see if it will work in anlz_fibmatrix  - pretend it's fecal coliform

```{r}
fake_fib <- fibdat |> 
    rename(epchc_station = station,
           fcolif = ecocci)
anlz_fibmatrix(fake_fib)
show_fibmatrix(fake_fib, stas = unique(fake_fib$epchc_station),
               yrrng = c(2018, 2022),
               lagyr = 1)
# not pretty, but it does work
```




`anlz_fibmap()`  - categorize based on thresholds  

`anlz_fibmatrix()`  

## Entire data frame  

```{r}
anlz_enteromatrix(fibdat, lagyr = 1)
show_enteromatrix(fibdat, lagyr = 1)

anlz_enteromap(fibdat)


# playing with maps through time
show_enteromap(fibdat, yrsel = 2020, mosel = 8)
show_enteromap(fibdat, yrsel = 2020, mosel = 9)
show_enteromap(fibdat, yrsel = 2020, mosel = 10)
show_enteromap(fibdat, yrsel = 2020, mosel = 11)
show_enteromap(fibdat, yrsel = 2020, mosel = 12)
show_enteromap(fibdat, yrsel = 2021, mosel = 1)
show_enteromap(fibdat, yrsel = 2021, mosel = 2)
show_enteromap(fibdat, yrsel = 2021, mosel = 3)
show_enteromap(fibdat, yrsel = 2021, mosel = 4)
show_enteromap(fibdat, yrsel = 2021, mosel = 5)
```

maybe something we could do would be change shape based on whether the sample was wet or dry.....

## Subsets of wet and dry  

### Wet  

```{r}
anlz_enteromatrix(fibwet, lagyr = 1)
show_enteromatrix(fibwet, lagyr = 1)

anlz_enteromap(fibwet)
```

### Dry  

```{r}
anlz_enteromatrix(fibdry, lagyr = 1)
show_enteromatrix(fibdry, lagyr = 1)

anlz_enteromap(fibdry)
```

Weird; I'd have thought we'd see more 'bad' colors in the wet subset. Maybe we're seeing more in the dry just because there are more samples, period???  

In the wet samples, the 'bad' scores seem to be more associated with certain years (granted this is a small sample size), whereas in the dry samples, it's more that certain stations seem more problematic than others.  


# Generate Report Card tables/graphics  

`show_fibmatrix()`  

`show_thrplot()`  

`show_boxplot()`  





