---
title: "Untitled"
format: html
---

```{r}
library(dataRetrieval)
```


```{r}
# pulled these from 'StationNam' field of catchments shapefile
# some rows had multiple stations combined with '+'
# I separated those out here
stns <- c("21FLHILL_WQX-101",
          "21FLHILL_WQX-102",
          "21FLHILL_WQX-103",
          "21FLHILL_WQX-619",
          "21FLHILL_WQX-104",
          "21FLHILL_WQX-501",
          "21FLHILL_WQX-502",
          "21FLHILL_WQX-504",
          "21FLPDEM_WQX-05-06",
          "21FLPDEM_WQX-12-02",
          "21FLPDEM_WQX-19-02",
          "21FLPDEM_WQX-22-01",
          "21FLPDEM_WQX-24-02",
          "21FLTPA_WQX-G1SW0095",
          "21FLTPA_WQX-G1SW0099",
          "21FLHILL_WQX-74",
          "21FLHILL_WQX-109",
          "21FLHILL_WQX-133",
          "21FLHILL_WQX-144",
          "21FLHILL_WQX-560",
          "21FLHILL_WQX-561",
          "21FLTPA_WQX-G1SW0093",
          "21FLHILL_WQX-594",
          "21FLHILL_WQX-595",
          "21FLHILL_WQX-596",
          "21FLHILL_WQX-597",
          "21FLHILL_WQX-598",
          "21FLHILL_WQX-599",
          "21FLHILL_WQX-2",
          "21FLCOSP_WQX-32-03",
          "21FLCOSP_WQX-44-02",
          "21FLCOSP_WQX-SC-01",
          "21FLHILL_WQX-136",
          "21FLHILL_WQX-264",
          "21FLHILL_WQX-592",
          "21FLTPA_WQX-G1SW0092",
          "21FLHILL_WQX-112",
          "21FLHILL_WQX-180",
          "21FLDOH_WQX-MANATEE152",
          "21FLDOH_WQX-MANATEE153",
          "21FLHILL_WQX-22",
          "21FLMANA_WQX-FC1",
          "21FLCOSP_WQX-48-03",
          "21FLCOSP_WQX-COSP580",
          "21FLCOSP_WQX-NORTH CANAL",
          "21FLCOSP_WQX-SOUTH CANAL",
          "21FLCOSP_WQX-CENTRAL CANAL",
          "21FLPDEM_WQX-39-01",
          "21FLPDEM_WQX-39-05",
          "21FLMANA_WQX-LM3",
          "21FLMANA_WQX-LM4",
          "21FLMANA_WQX-LM5",
          "21FLMANA_WQX-LM6")

```

Characteristic Names Enterococci and Enterococcus; siteIDs only trying 3

https://www.waterqualitydata.us/#siteid=21FLCOSP_WQX-CENTRAL%20CANAL&siteid=21FLCOSP_WQX-NORTH%20CANAL&characteristicName=Enterococci&characteristicName=Enterococcus&mimeType=csv&providers=NWIS&providers=STORET

```{r}
# set up the query
# can enter start dates here if desired
# (into the args list)
entero_names <- c("Enterococci",
                  "Enterococcus")
siteid <- stns

startDateLo <- as.Date("2012-01-01")
# startDateHi would be an end date

args <- list(
    siteid = siteid,
    characteristicName = entero_names,
    startDateLo = startDateLo
)
```



See how many records will be returned:  

```{r}
wqp_summ <- readWQPdata(args, querySummary = TRUE)

n_results <- wqp_summ$`total-result-count`
n_sites <- wqp_summ$`total-site-count`
```

The query will return **`r n_results`** results from **`r n_sites`** stations.  

And get the data:

```{r}
wqp_dat <- readWQPdata(args, querySummary = FALSE)

# the following actually returns more columns (80 rather than 67)
wqp_narrow <- readWQPdata(args, querySummary = FALSE, dataProfile = "narrowResult")
```

Fields of interest are:  

OrganizationIdentifier  
OrganizationFormalName  
MonitoringLocationIdentifier = station  
ActivityTypeCode (e.g. "Sample-Routine")  

ResultMeasureValue - the result (has characters in here too - "Not Reported")  
ResultMeasure.MeasureUnitCode  
ActivityStartDate
ActivityStartTime.Time (local time)  
ActivityStartTime.TimeZoneCode  

[ActivityStartDateTime is in UTC]  

ResultAnalyticalMethod.MethodIdentifier - in case you want to include/exclude specific methods  
DetectionQuantitationLimitMeasure.MeasureValue  
ResultLaboratoryCommentText  
